---
date: 04-12-2019
output:
  pdf_document:
    includes:
      in_header: header.tex
---

\tableofcontents
\newpage

```{r, options_chunk, include=FALSE}
knitr::opts_chunk$set(
    cache = FALSE, eval = TRUE, echo = FALSE, fig.align = "center",
    fig.width = 12, fig.height = 8, warning = FALSE, message = FALSE
)
source("~/github/pecs/src/r/master.R")
```

\section{Trends in the data}

\subsection{What's the grand picture?}

```{r, detrendAnalysis, include = FALSE}
source(file.path(path_project, "src", "r", "salvage", "01_detrended_analysis.R"))
```

Turnout has been declining for decades. Starting in 1970, participation in general
elections decreased from an average
`r sprintf("%.2f", 100 * mean(tillman$turnout[tillman$year == 1970]))` per cent to
`r sprintf("%.2f", 100 * mean(tillman$turnout[tillman$year == 2011]))` per cent in
2011. The figure above shows the development for all democracies in the data. With
the exceptions of Belgium and Denmark turnout has declined across the board
although differences in speed, magnitude, and monotonicity of the decline
are appearent.

\begin{figure}[ht]
    \caption{Trends in Turnout}
    \includegraphics[width=\textwidth]{~/github/pecs/out/scatter_smoother_detrendTurnout.pdf}

    \textbf{Note}: Confidence envelope at the .95 level.
\end{figure}

Importantly, trends that parallel turnout plague Tillman's two key predictors:
the formation of pre-electoral coalitions and their voting percentage. The
figure below shows Z-standardized versions of the fixed-effect transformed
response and its two key predictors. Although developments may not always align
to 100 percent, e.g., Italy, cases like Germany, the Netherlands, Greece or
Denmark highlight the potential for confounding trends in Tillman's analysis
clear.

\begin{figure}[ht]
    \caption{Shared Trends in Treatments and Response}
    \includegraphics[width=\textwidth]{~/github/pecs/out/line_commonTrend_Pec1_Turnout.pdf}

    \small{\textbf{Note}: Rug plots indicate election years.}
\end{figure}

\newpage


## What do the detrended results look like?

Several options to detrend panel data are on the table.
\begin{enumerate}
    \item Indicator variables for observations can account for all apparent
        trends in the data. However, they reduce statistical power quite a bit.
    \item Reduced form trend models of observation time, e.g., polynomials or
        splines, are easily implemented, but quickly overfit the data
    \item Sophisticated error structures, e.g., AR(1), correct standard
        errors and p-values. Unfortunately, they ignore the biased effect
        estimate.
    \item A set of lagged responses removes any trend up to the cardinality of
        the set, but it will bias results if the trend is more complex than
        that.
    \item Substantive models of the trend in the outcome. However, if those
        models are to reduce confoundedness then they should yield leverage on
        the treatment(s) as well.
\end{enumerate}

Keeping in mind our paper's purpose -- a replication and data feature --,
options one and two strike as most promising. Table
\ref{tbl:DetrendedRegressionResults} replicates Tillman's entire analysis using
cubic base [splines](https://en.wikipedia.org/wiki/B-spline) with three
knots.\footnote{
    The last report already presented the dummy variable approach.
}
The table reports panel robust standard errors, clustered by country. There are
three take-away messages.

```{r, results = "asis"}
texreg(fitted_models,
    digits = 3,
    override.se = lapply(bk_se, function(x){sqrt(diag(x))}),
    override.p = bk_pval,
    custom.coef.names = c("PEC", "Parties", "Disproportionality", "PR",
        "Plurality", "Closeness", "Economic Growth", "log(Income)", "Spline 1",
        "Spline 2", "Spline 3", "Spline 4",
        "PR $\\times$ Disproprtionality", "Plurality $\\times$ Closeness",
        "PEC $>$ 20\\%", "Vote PEC", "Small PEC", "Large PEC"
    ),
    reorder.coef = c(1, 15, 16, 17, 18, 2:14),
    caption = "Detrended Regression Results", caption.above = TRUE,
    label = "tbl:DetrendedRegressionResults"
)
```

\begin{enumerate}
    \item As before, the \textit{formation} of pre-electoral coalitions does
        not boost turnout once trends are taken into account. The estimated
        effect on PEC approaches zero and more sophisticated trend
        formulations will not substantively change this outcome: Shared trends
        confound Tillman's pivotal analysis.
    \item Interestingly, Models 2 through 4 imply that strong pre-electoral
        coalitions still boost turnout. According to Model 3, every additional
        vote percentage point which goes to PECS increases turnout by $0.027$
        percentage points. Moreover, Models 2 and 4 strongly suggest that this
        effect turns strong pre-electoral coalitions, i.e. PECs that account
        for at least 20 percent of the vote.

        However, voting results for PECs are downstream from turnout.
        Therefore, it is very likely that something else but time confounds
        these effects. Suspects range from institutions and party system
        polarization to observable properties of PECS, e.g. former cabinet
        status of its members or coalition type.
    \item Once the analysis has been detrended, many of the remaining
        regressors tend towards zero. At least, most of them do not flip sign.
        Consequently, not everything about Tillman's analysis is confounded by
        trends. There is something to be salvaged.
\end{enumerate}

\section{Rethinking institutional controls}

\section{Bringing polarization back in}
