---
title: "Salvage"
author: "Dag Tanneberg"
output: pdf_document
---

```{r, options_chunk, include=FALSE}
knitr::opts_chunk$set(
    cache = FALSE, eval = TRUE, echo = FALSE, fig.align = "center",
    fig.width = 12, fig.height = 8, warning = FALSE, message = FALSE
)
source("~/github/pecs/src/r/master.R")
```

# Trends in the data

## What's the grand picture?
```{r, detrendAnalysis, include = FALSE}
source(file.path(path_project, "src", "r", "salvage", "01_detrended_analysis.R"))
```

![Trends in Turnout](~/github/pecs/out/scatter_smoother_detrendTurnout.pdf)

Turnout has been declining for decades. Starting in 1970, participation in general 
elections decreased from an average
`r sprintf("%.2f", 100 * mean(tillman$turnout[tillman$year == 1970]))` per cent to
`r sprintf("%.2f", 100 * mean(tillman$turnout[tillman$year == 2011]))` per cent in 
2011. The figure above shows the development for all democracies in the data. With 
the exceptions of Belgium and Denmark turnout has declined across the board
although differences in speed, magnitude, and monotonicity of the decline
are appearent.

Importantly, trends that parallel turnout plague Tillman's two key predictors: the
formation of pre-electoral coalitions and their voting percentage. The figure
below shows Z-standardized versions of the fixed-effect transformed response and
its two key predictors. Although developments may not always align to 100 percent,
e.g., Italy, cases like Germany, the Netherlands, Greece or Denmark make the
potential for confounding trends in Tillman's analysis abundantly clear.

![Shared Trends in Treatment and Response](~/github/pecs/out/line_commonTrend_Pec1_Turnout.pdf)

\newpage


## What do the detrended results look like?

```{r, results = "asis"}
texreg(fitted_models,
    digits = 3,
    override.se = lapply(bk_se, function(x){sqrt(diag(x))}),
    override.p = lapply(bk_se, function(x){sqrt(diag(x))}),
    custom.coef.names = c("PEC", "Parties", "Disproportionality", "PR",
        "Plurality", "Closeness", "Economic Growth", "Income", "Spline 1",
        "Spline 2", "Spline 3", "Spline 4",
        "PR $\\times$ Disproprtionality", "Plurality $\\times$ Closeness",
        "PEC $>$ 20\\%", "Vote PEC", "Small PEC", "Large PEC"
    ),
    reorder.coef = c(1, 15, 16, 17, 18, 2:14)
)
```


