---
output:
    pdf_document:
        includes:
            in_header: header.tex
    documentclass: article
---

\markboth{GANGHOF, EPPNER, AND TANNEBERG}{Turnout and Pre-Electoral Coalitions}

```{r, optionsChunk, include = FALSE}
knitr::opts_chunk$set(
    cache = FALSE, eval = TRUE, echo = FALSE, fig.align = "center",
    fig.width = 12, fig.height = 8, warning = FALSE, message = FALSE,
    fig.lp="fig:", fig.pos="t"
)
source("~/github/pecs/src/r/master.R")
```

\maketitle

\section{Introduction}

What explains turnout in general elections with high numbers of competing parties?
After all, as the number of political parties increases coalition cabinets become
more likely whose format usually depends exclusively on the bargains struck in
parliament. ``The presence of many parties may mean that voters have little say in
the actual selection of the government'' \cite[118]{Blais.2006}, effectively
reducing the incentive to turn out and vote. Extant research suggests that
pre-electoral coalitions mitigate the adverse effect of multipartism as they
provide everyone with additional information on credible coalition options
\cite[197]{Smith.2018}. However, prior empirical tests of that argument
suffer from two non-negligble constraints. First, results are based on a small set
of countries and a limited number of elections. As a consequence, their
generalizability is questionable. Second, and more importantly, prior research
frequently aggregates data to the election level, disregarding information on
different types, electoral strength or prior incumbency of pre-electoral
coalitions. This paper presents new data on pre-electoral coalitions that have
higher temporal and spatial coverage and that allow closer scrutiny of each
covered pre-electoral coalition. The data highlight substantial differences
between pre-electoral coalitions and imply that only coalitions which formulate
a clear alternative to the incumbent government increase turnout in multiparty
elections.

\section{Why does multipartism decrease turnout?}

\section{What pre-electoral coalitions increase government identifiability?}

\section{Results}

\subsection{Replication}

\cite{Tillman.2015} relies on a standard fixed-effects panel design in order to
study the effect of pre-electoral coalitions on turnout in 243 parliamentary
elections between 1970 and 2011. In the causal craze of this day and age
fixed-effects designs have become the "gold standard" \cite[139]{Bell.2015} in
many fields because they eliminate heterogeneity bias. More precisely, by way of
simple transformations of the data fixed-effects designs remove time-invariant
unit-specific unobservables \cite[508]{Halaby.2004}. In Tillman's case that
refers to permanent ideosyncracies of the twenty parliamentary democracies in
his data. Having eliminated unobserved heterogeneity among the units,
fixed-effect designs return unbiased estimators of a hypothesized causal effect
as long as the disturbances are mean-independent of the independent variable(s)
across \textit{all} measurement occasions \cite[482]{Wooldridge.2009}. Despite
their appeal for causal inference, however, fixed-effects designs have recently
come under attack for various reasons [add references]. A most recent
Monte-Carlo study even advises advise scholars to distrust fixed-effects
estimators whenever there is a reason to suspect "the presence of any dynamic
misspecification or omitted time-varying variables"
\cite[26; our emphasis]{Pluemper.2018}. Evidence of such misspecification in
Tillman's analysis will be presented in the following.

Misspecifications of fixed-effects designs arise in numerous ways of which
failure to account for non-stationarity is particularly troublesome.\footnote{
    This paragraph summarizes what I understand so far. I am pretty sure,
    there is room for improvement. Hints to easy-going literature on time
    series and panel data analysis are very welcome.
}
Stationarity requires that the mean, variance, and autocorrelation of a
time-series are finite and constant over time \cite[378]{Wooldridge}. Its
violations take the form of either mean shifts or trends. While we have little
reason to suspect the former in Tillman's data, we do find evidence for the
latter, leading us to argue that Tillman's results suffer from the so-called
"spurious" or "nonsense regression problem" \cite{Granger.1974,Wooldridge.2009}.
To elaborate, if trends mar the data, then otherwise independent variables may
still seem strongly related. The reasons are twofold. First, trends go hand in
hand with serial or autocorrelation among the errors. Observations close in time
tend to be more similar than others \cite[432]{Fox.2011}. As a consequence, the
usual test statistics and any tests of hypothesis based on them are invalid.
Second, the estimated effect itself may be biased because the trend -- more
precisely, its (unobserved) causes -- confounds the relationship between two
variables. In other words, two trending variables may seem related because they
share a time-varying, common cause. Progressing from simple visual inspection to
standard diagnostics and detrending tools, we demonstrate below that a common
trend in turnout and the formation of pre-electoral coalitions biases
much, but possibly not not all of Tillman's results.

```{r, edaNonStationarity, fig.cap="Visual Assessment of Stationarity"}
rm(list = ls()[!(ls() %in% clean_workspace)])
pdta <- filter(country_panel, in_tillman == 1) %>%
    select(iso3c, year, turnout, pec1, vote_pec) %>%
    mutate(turnout = turnout / 100, vote_pec = vote_pec / 100)
tmp <- aggregate(
    pdta[, c("turnout", "pec1", "vote_pec")],
    list(year = pdta[["year"]]),
    FUN = mean, na.rm = TRUE
)
tmp[, "statistic"] <- "mu"
tmp2 <- aggregate(
    pdta[, c("turnout", "pec1", "vote_pec")],
    list(year = pdta[["year"]]),
    FUN = sd, na.rm = TRUE
)
tmp2[, "statistic"] <- "sigma"
pdta <- bind_rows(tmp, tmp2) %>%
    gather("key", "value", turnout, pec1, vote_pec) %>%
    mutate(
        key_labels = factor(
            key, levels = c("turnout", "pec1", "vote_pec"),
            labels = c("Turnout", "PEC contested", "Vote PEC")
        ),
        statistic_labels = factor(statistic, levels = c("mu", "sigma"),
                labels = c("Mean", "Std. Dev.")
            )
    )
ggplot(data = pdta, aes(x = year, y = value)) +
    geom_point(shape = 21, fill = "white") +
    geom_smooth(method = "lm", se = FALSE, formula = y ~ poly(x, 3), colour = "black", size = .7) +
    facet_grid(statistic_labels ~ key_labels) +
    labs(caption = "Note: Features rescaled to unit-ranges. Cubic polynomials added."
    ) +
    ggthemes::theme_fivethirtyeight()
```

The properties of stationary data imply graphical tests, the simplest of
which are scatterplots of means and variances against time. If the data on
turnout and pre-electoral coalitions are stationary, then their mean and
variance should fluctuate only randomly over time. Figure
\ref{fig:edaNonStationarity}, however, implies the opposite. It plots annual
means (top row) and standard deviations (bottom row) for each of Tillman's
pivotal variables.\footnote{
    The publication features a number of indicator variables which discretize
    the electoral strength of pre-electoral coalitions. For example, one binary
    indicator takes the value of 1 whenever at least one pre-electoral
    coalition received at least 20 percent of the national vote
    \cite[731]{Tillman}. As these binary indicators are functions of the
    overall electoral strength of pre-electoral coalitions, they will
    follow the same trend, if any.
}
Cubic polynomials have been added to each plot and highlight the development
of their central tendency. All in all, Figure \ref{fig:edaNonStationarity} graph
casts severe doubt on the stationarity of the data.

\begin{enumerate}
    \item As documented by prior research, average turnout has been declining
        for years \cite{Hooghe.2017}. Starting at 88 percent in 1970, the
        annual mean in the top-left panel reaches an all-time low at 66 percent
        in 2000 and shows no sign of reversion until the end of
        the observation period. Pre-electoral coalitions exhibit a more
        complex, though not unrelated trend. As shown in the middling
        panel of the top row, PECs become more prevalent in national elections
        until the mid-1980s. Later on, their attraction nosedives -- in parallel
        to turnout. Voting results for PECs, in contrasts, do not exhibit such
        clear-cut patterns.
    \item Turning to their second moment, evidence against stationarity of the
        data is not as obvious. On the one hand, the standard deviation 
        about the annual average in turnout has monotonically increased over
        time. This observation clearly speaks to non-stationarity. On the other
        hand, the variances in the formation and showing of pre-electoral
        coalitions are prone to some notable outliers that may skew the overall
        impression from the diagram.
    \item Finally, Figure \ref{fig:edaCommonTrend} in the Appendix
        disaggregates the data to the country-level. As can be seen from the
        individual panels, all three indicators, turnout, the formation of
        pre-electoral coalitions, and their voting result, seem to change
        systematically over time. However, there are stark differences between
        countries which occasionally exhibit very complex trends.
\end{enumerate}

As noted repeatedly above, visual assessments of statistical phenomena are
informative but often subjective. Hence, they should be accompanied by more
formal evidence whenever possible. In the case of non-stationarity, tests for
autocorrelation among the residuals constitute one option. Table
\ref{tbl:TestsSerialCorrelation} reports several different autocorrelation
diagnostics.\footnote{
    Additional unit root tests would be awesome, but package plm fails to
    compute. Advise is welcome.
}
All of them all evaluate Tillman's pivotal regression models \cite[733, Table 4;
Models 2 and 4]{Tillman.2015}. The Durbin–Watson statistic detects first-order
autocorrelation among the residuals. Its classical form does not account for
panel-data and has thus been generalized by \cite{Baltagi.1999}. If first-order
autocorrelation (AR(1)) was not an issue, than either variant of the statistic
should approach 2. This is clearly not the case. Rather, at values about 1
either variant of the Durbin-Watson statistic implies positive autocorrelation
among the residuals. In contrast to other tests for autocorrelation, the
Breusch–Godfrey test allows the user to choose the order of autocorrelation to
test for. The table shows evidence in favor of AR(1) and AR(2) processes. At 1
respectively 2 degrees of freedom all 4 $\chi^2$-statistics reach very highly
levels of statistical significance and reject the null of no-autocorrelation.
Finally, \cite[Sec. 10.5.4]{Wooldridge.2010} proposes a test for serial
correlation among the errors in fixed–effects panel models. Table
\ref{tbl:TestsSerialCorrelation} reports the corresponding F-scores and degrees
of freedom. Either test is highly statistically significant and rejects the null
of no-serial correlation. In conclusion, the table supports our earlier
impression of non-stationary in the data and sheds further doubt on Tillman's
results.


```{r, reportSerialCorrelation, eval = FALSE}
data_to_fit <- filter(
    country_panel, in_tillman == 1 & !(iso3c %in% c("LUX", "CAN"))
)
data_to_fit <- pdata.frame(data_to_fit, index = c("iso3c", "year2"))
fit_contested <- plm(
    turnout ~ pec1 + enep + disprop * pr + plurality * closeness + growth +
    lnincome, model = "within", data = data_to_fit
)
fit_result <- plm(
    turnout ~ vote_pec + enep + disprop * pr + plurality * closeness + growth +
    lnincome, model = "within", data = data_to_fit
)
vapply(list("contested" = fit_contested, "result" = fit_result), function(x) {
    pdw <- pdwtest(x)[["statistic"]]
    pbn <- pbnftest(x)[["statistic"]]
    pbg1 <- pbgtest(x, order = 1)[["statistic"]]
    pbg1_p <- pbgtest(x, order = 1)[["p.value"]]
    pbg2 <- pbgtest(x, order = 2)[["statistic"]]
    pbg2_p <- pbgtest(x, order = 2)[["p.value"]]
    out <- c(pdw, pbn, pbg1, pbg1_p, pbg2, pbg2_p)
    return(out)
},
    FUN.VALUE = numeric(6)
)
pwartest(fit_contested)
pwartest(fit_result)

pbgtest(fit_contested, order = 2)
```

\begin{table}
    \centering
    \begin{tabular}{lcc}
        \toprule
        ~ & \multicolumn{2}{c}{Test Statistic} \\
            & PEC contested & PEC result \\
        \midrule
        Classical Durbin–Watson & 1.2 & 1.2 \\
        Modified Durbin–Watson & 0.9 & 0.9 \\
        Breusch-Godfrey $\chi^2_{AR(1)}$ & 48.1 & 44.7 \\
        Breusch-Godfrey $\chi^2_{AR(2)}$ & 48.2 & 44.8 \\
        Wooldridge & $F_{(1, 220)} = 18.31$ & $F_{(1, 216)} = 17.96$ \\
        \bottomrule
    \end{tabular}
    \caption{Tests for Serial Correlation among the Residuals}
    \label{tbl:TestsSerialCorrelation}
\end{table}


How to deal with non-stationarity in fixed-effects designs? One solution
explicitly includes a trend component into the analysis. There are several
options to do so. Firstly, a two-way error components model eliminates
unobserved heterogeneity among the units \text{and} the time points
\cite[982]{Bruederl.2010}. It is akin to a unit-fixed-effects model that
includes also time dummies and constitutes a conventional choice in panel data
analysis. Secondly, a time index component can be included as an independent
variable. The index variable reduces confounding due to trends in the data. A
somewhat more flexible version of this approach interacts the time index with
the unit-fixed effects to estimate unit-specific trends (growth curves).
A third solution to the problem of non-stationarity transforms the data,
e.g., by taking first-differences or applying a Prais-Winston
transformation\footnote{
    We did not get around to implement the Prais-Winston regression yet.
}.
Finally, dynamic-panel models have been suggested as a fix for serial
correlation among the errors [REFERENCE]. The perhaps best known variant of
this approach goes back to \cite{Beck.1995}. It adds a one-period lag of
the dependent variable to the predictors and uses panel-corrected standard
errors. Each of these attempted solutions approaches non-stationarity from a
different angle, and, taken together, they add up to a rigorous reassessment of
the relationship between pre-electoral coalitions and turnout in parliamentary
democracies.

We present results for each of the previously described variations\footnote{
    Again, we didn't get around to do Prais-Winston, yet.
}
on the fixed-effects scheme in Figure \ref{fig:reportDetrendedResults}. The top
row plots the marginal effect of PEC formation on turnout, while the bottom
row shows the marginal effect of the total PEC vote share on turnout.\footnote{
    Following the original publication, turnout and Vote PEC are measured in
    percent.
}
Confidence intervals at the 90 and 95 percent levels have been added to each
panel.\footnote{
    All models use panel-corrected standard errors, clustered by country.
}

The left-most column replicates Tillman's original analysis which does not
account for non-stationarity. It implies that pre-electoral coalitions boost
turnout (top) and that an increase in their share of votes concurs with an
increase in turnout (bottom). All confidence intervals securely exclude $0$,
meeting conventional standards of statistical significance. Things change
markedly once non-stationarity is taken into account. Going from left to right
in the top row, the mobilizing potential of PEC entering elections is (a)
immediately and notably reduced in magnitude, and (b) almost consistently 
indistinguishable from $0$. The dynamic panel model constitutes the sole
exception to this rule, but it should not receive too much emphasis. After all,
even this modeling approach matches the more general pattern, it primarily
addresses serial correlation among the errors, and its confidence intervals just
barely exclude $0$. 

Results in the bottom row are more mixed. In all but two instances stronger
electoral showings of PECs coincide with higher turnout even after
non-stationarity has been accounted for and even though the magnitude of the
estimated coefficient declines by up to 50 percent. Yet, many estimates still
meet conventional levels of statistical significance. It is noteworthy, though,
that the unit-specific trend and the first-difference formulation render the
electoral strength of PECs statistically indistinguishable from $0$. These
models apply the most stringent fixes of non-stationarity, and they clearly
reduce first order autocorrelation among the among the errors (see
\ref{tbl:ReplicationResults}). In short, once non-stationarity is taken into
account scant evidence supports the mobilizing effect of PEC formation, but the
data hint to a continued association between their electoral strength and
turnout. This hint will be subject to analysis in the following section.

```{r, reportDetrendedResults, fig.cap="Estimated Effect of PECs on Turnout"}
# Initialize
rm(list = ls()[!(ls() %in% clean_workspace)])
packages <- c("plm", "sandwich", "lmtest", "texreg")
for(i in packages){
    if(!(i %in% rownames(installed.packages()))) {
        cat("Now installing required package:\t", i)
        install.packages(i, dependencies = TRUE)
    }
    library(i, character.only = TRUE)
}


# Declare Functions
extract_results <- function (model, of.interest = NULL, pcse = "group") {
    out <- cbind(beta = coef(model), se = sqrt(diag(vcov(model))))
    if(!is.null(pcse)) {
        out[, "se"] <- sqrt(diag(plm::vcovBK(model, cluster = pcse)))
    }
    if (!is.null(of.interest)) {
        out <- out[grep(of.interest, rownames(out), perl=TRUE), ]
    }
    return(out)
}


# Declare string constants
panel_id <- c("country", "year2")
control <- paste(
    c("enep", "disprop * pr", "plurality * closeness", "growth", "lnincome"),
    collapse = " + "
)
treatments <- c("PEC contested" = "pec1", "Vote PEC" = "vote_pec")
response <- "turnout"
trends <- c("None"="None", "Two-way Error\nComponent"="twoways",
    "Common\nTrend"="t", "Unit-specific\nTrend"="t*iso3c",
    "First\nDifference"="fd", "Lagged DV"="turnout_l1"
)


# Prepare data
data_to_fit <- filter(country_panel, (in_tillman == 1) &
    !(iso3c %in% c("LUX", "CAN"))
) %>%
    group_by(iso3c) %>%
    mutate(turnout_l1 = dplyr::lag(turnout, n = 1, order_by = year2)) %>%
    ungroup()
data_to_fit <- pdata.frame(data_to_fit, index = c("iso3c", "year2"))


# Fit models
fitted_models <- list()
for (trend in trends){
    for(treatment in treatments) {
        assign(
            "frm",
            paste0(response, " ~ ", paste(treatment, control, trend, sep=" + "))
        )
        if (trend == "None") {
            assign(
                "frm",
                paste0(response, " ~ ", paste(treatment, control, sep=" + "))
            )
        }
        if (trend == c("twoways")) {
            assign(
                "frm",
                paste0(response, " ~ ", paste(treatment, control, sep=" + "))
            )
            fitted_models[[paste(trend, treatment, sep = "_")]] <- plm(
                as.formula(frm), effect="twoways", model="within", data=data_to_fit
            )
            next
        }
        if (trend == c("fd")) {
            assign(
                "frm",
                paste0(response, " ~ ", paste(treatment, control, sep = " + "))
            )
            fitted_models[[paste(trend, treatment, sep = "_")]] <- plm(
                as.formula(frm), effect="individual", model="fd", data=data_to_fit
            )
            next
        }
        fitted_models[[paste(trend, treatment, sep = "_")]] <- plm(
            as.formula(frm), model = "within", data = data_to_fit
        )
    }
}


# Generate coefficient plot
to_plot <- vapply(fitted_models[c(1:8, 11:12)],
    FUN=extract_results, FUN.VALUE=numeric(2), of.interest="pec"
)
to_plot <- cbind(
    to_plot, vapply(fitted_models[c("fd_pec1", "fd_vote_pec")],
    FUN = function(model){
        out <- cbind(beta = coef(model), se = sqrt(diag(vcov(model))))
        out <- out[grep("pec", rownames(out), perl=TRUE), ]
        return(out)
    },
    FUN.VALUE = numeric(2)
    )
)
to_plot <- t(to_plot)
treatment_locations <- str_locate(rownames(to_plot), "vote_pec$|pec1$")
treatment_labels <- str_sub(rownames(to_plot), treatment_locations)
treatment_labels <- factor(treatment_labels, treatments, names(treatments))
trend_labels <- str_split_fixed(rownames(to_plot), "vote_pec|pec1", 2)[, 1]
trend_labels <- str_sub(trend_labels, 1, str_locate(trend_labels, "_$")[, 1]-1)
trend_labels <- factor(trend_labels, trends, names(trends))
z <- qnorm(c(.1, .05) / 2, lower = FALSE)
pdta <- tibble(beta=to_plot[, "beta"], se=to_plot[, "se"],
    trend_label=trend_labels, treatment_label=treatment_labels
) %>%
    mutate(lower90 = beta - z[1] * se, upper90 = beta + z[1] * se) %>%
    mutate(lower95 = beta - z[2] * se, upper95 = beta + z[2] * se)
ggplot(data = pdta, aes(x = trend_label, y = beta)) +
    geom_linerange(aes(ymin = lower95, ymax = upper95)) +
    geom_linerange(aes(ymin = lower90, ymax = upper90), size=1.1) +
    geom_point(size = 2, shape = 21, fill = "white") +
    labs(x = "Correction for Non-Stationarity", y = "Effect Estimate",
        caption = "Note: CIs at 90 and 95 percent."
    ) +
    facet_grid(treatment_label ~ ., scales = "free") +
    ggthemes::theme_fivethirtyeight() +
    theme(axis.title = element_text())


# Return BTON
ar_statistics <- vapply(fitted_models,
    function(m) {
        pdw <- pdwtest(m)[["statistic"]]
        pbg1 <- pbgtest(m, order = 1)[["statistic"]]
        pbg2 <- pbgtest(m, order = 2)[["statistic"]]
        c(pdw, pbg1, pbg2)
    },
    FUN.VALUE = numeric(3)
)
ses <- lapply(fitted_models[c(1:8, 11:12)], vcovBK, cluster = "group")
ses[["fd_pec1"]] <- vcov(fitted_models[["fd_pec1"]])
ses[["fd_vote_pec"]] <- vcov(fitted_models[["fd_vote_pec"]])
ses <- ses[names(fitted_models)]
pvals <- lapply(names(fitted_models),
    function(x) coeftest(fitted_models[[x]], vcov = ses[[x]])[, 4]
)
names(pvals) <- names(fitted_models)
texreg(
    fitted_models,
    file = file.path(path_project, "artcl", "detrended_models.tex"),
    digits = 2,
    omit.coef = "(t:)",
    override.se = lapply(ses, function(x){sqrt(diag(x))}),
    override.p = pvals,
    custom.coef.names = c("PEC contested", "Parties", "Disproportionality",
        "PR", "Plurality", "Closeness", "Economic Growth", "log(Income)",
        "Disproportionality $\\times$ PR",  "Plurality $\\times$ Closeness",
        "Vote PEC", "$t$", "Constant", "Turnout$_{t-1}$"
    ),
    reorder.coef = c(1, 11, 2:10, 12, 14, 13),
    booktabs = TRUE, sideways = TRUE,
    stars = c(0.001,  0.01, 0.05, 0.1),
    caption = "Replication Results",
    label = "tbl:ReplicationResults",
    custom.note = paste0("Panel corrected SEs in parantheses ",
        "\\cite{Beck.1995}, clustered by country. %stars."
    )
)
# Manual edits to BTON
tmp <- readLines(file.path(path_project, "artcl", "detrended_models.tex"))
tmp <- tmp[!(tmp == "")]  # drop empty strings
tmp <- tmp[!grepl("\\usepackage", tmp)]  # drop usepackage entries
tmp[grepl("[a-z]\\_[a-z]", tmp)] <- paste(
    "Detrending Technique", "\\multicolumn{2}{c}{None}",
    "\\multicolumn{2}{c}{2way EC}",
    "\\multicolumn{2}{c}{Common Trend}",
    "\\multicolumn{2}{c}{Unit-specific Trend}",
    "\\multicolumn{2}{c}{FD}",
    "\\multicolumn{2}{c}{LDV} \\\\",
    sep = " & "
)  # replace table header
split_at <- which(tmp == "\\bottomrule")
ar_labels <- c("Durbin-Watson",
  "Breusch-Godfrey $\\chi^2_{AR(1)}$", "Breusch-Godfrey $\\chi^2_{AR(2)}$"
)
lines_to_add <- vector("character", length(ar_labels))
for(i in seq(lines_to_add)) {
    lines_to_add[i] <- paste0(
        paste0(ar_labels[i], " & "),
        paste(signif(ar_statistics[i, ], 3), collapse = " & "),
        "\\\\"
    )
}
tmp <- c(tmp[1:(split_at - 1)], lines_to_add, tmp[split_at:length(tmp)])
writeLines(tmp, con = file.path(path_project, "artcl", "detrended_models.tex"))


# housekeeping
for (p in packages) detach(paste("package", p, sep = ":"), character.only = TRUE)
rm(list = ls()[!(ls() %in% clean_workspace)])
```

\subsection{Exploratory Analysis: What Pre-electoral Coalitions matter?}

The previous section casts doubt on the mobilizing potential of
pre-electoral coalitions, but it does not prove the suspicion entirely unfounded.
Participation in elections does not increase when PECs enter the race, but their
strong showing tends to coincide with higher levels of turnout. Notwithstanding,
this seemingly robust association is theoretically shallow. If PECs provide
information on credible coalition governments and thereby incentivize turnout,
then we should focus on that message itself, not its ultimate effect. After all,
its ultimate effect may be endogenous to any number influences from the macro,
meso or micro level of parliamentary elections. The following section turns to a
number of PEC properties that may drive the seemingly robust association between
turnout and PEC total vote share. It concludes on a skeptical note and calls
for analyses with more granular data.

This exploratory endeavor takes advantage of new data on PECs and their
properties compiled by [GANGHOF AND EPPNER]. Following \cite{Golder.2006}, the
data define as pre-electoral coalition any group of parties that publicly and
mutually announces its intention to form or continue a coalition government
prior to election night. Any such statement shrinks the set of government
alternatives and thereby structures post-election coalition bargaining.
Notwithstanding, pre-electoral coalitions differ in their degree of commitment.
Some PECs merely announce their intention to cooperate come election dawn,
whereas others provide voters with detailed voting instructions, formulate joint
programs or field common candidates. Emphasizing those differences, the data
classify PECs by their type, the existence of a joint manifest, and their
incumbency status. For the purpose of this exploratory analysis, we presume that
such differences between pre-electoral coalitions reflect different levels of
sincerity among political parties and, thus, send more powerful messages to
voters.

```{r, ourPecsInTillman, eval=FALSE}
# Return totals for table
rm(list = ls()[!(ls() %in% clean_workspace)])
table_data <- filter(country_panel, (in_tillman == 1))
type_indicators <- grep("any_type[1-9]", names(table_data))
apply(table_data[, c("pec_neu", "any_incumbent", "any_prog")], 2, table)
apply(table_data[, type_indicators], 2, sum)
apply(table_data[table_data[, "any_incumbent"] == 1, type_indicators], 2, sum)
apply(table_data[table_data[, "any_prog"] == 1, type_indicators], 2, sum)
rm(list = ls()[!(ls() %in% clean_workspace)])
```

\begin{table}[ht]
    \centering
    \begin{tabular}{rccccccc}
        \toprule
        \multicolumn{2}{c}{} & \multicolumn{6}{c}{By Type} \\
        \cmidrule{3-8}
        ~ & Total & Public & Joint & Nomination & Vote Transfer & Dual-Ballot\\
        \multicolumn{1}{l}{Any PEC} & Elections & Commitment & List & Agreement & Instructions & Instructions & Other\\
        \midrule
        observed?           & 157 & 63 & 36 & 46 & 0 & 11 & 9\\
        incumbent?          &  46 & 31 & 21 &  3 & 0 &  6 & 1\\
        joint program?      &  64 & 32 & 14 & 32 & 0 &  1 & 3\\
        \bottomrule
        \begin{multicolumn}{8}{l}{\textit{Reading example}: The incumbent government formed a pre-electoral coalition in $46$ elections. Those
        elections coincided with $31$ elections during which at least one PEC
        was publicly announced. Elections may be double-counted.
    }
    \end{tabular}
    \caption{Absolute Frequency of PECs in 243 National Elections}
    \label{tbl:AbsFreqPecsInElections}
\end{table}

Table \ref{tbl:AbsFreqPecsInElections} summarizes our account of pre-electoral
coalitions for Tillman's 243 elections in 20 parliamentary democracies between
1970 and 2011.\footnote{
    Ganghof's and Eppner's data cover 563 elections in 35 democracies between
    1945 and 2015. In order to keep our results comparable
    with prior research, we limit ourselves to the set of parliamentary
    elections studied by Tillman. Tables \ref{tbl:summaryPecs} and
    \ref{tbl:SummaryPecProperties} in the Appendix summarize all the
    available data. Note, however, that there are occasionally non-trivial 
    differences between our data and what Tillman's data counts as PEC. For 
    instance, our data regards the decades-long cooperation between the German
    CDU and CSU as a pre-electoral whereas Tillman's data does not. As will
    become clear shortly, those difference do not interfere with the results.
}
According to our data, pre-electoral coalitions formed in 157, i.e. more than
half of all 243 elections. The incumbent government formed a PEC in roughly 1
out of 5 elections, and participants of at least one PEC agreed on a joint
program in approximately 25 percent of all elections. Mind the level of 
aggregation: Following Tillman's example, these figures reflect the election-year
level. Therefore, categories are not mutually exclusive and pre-electoral
coalitions may be double counted.\footnote{
    For instance, in 1983 the \textit{incumbent} German government, CDU/CSU and 
    FDP,  \textit{publicly announced} their intention to continue their
    coalition after the election. At the same time, CDU and CSU maintained
    a separate, non-incumbent, but \textit{nomination agreement}.
}
Regardless, public announcements of coalition intent are the modal form
of PEC across all elections. Joint lists and nomination agreements among
political parties come in second. When taken together, these categories may even 
outweigh public announcements. By comparison, vote transfer and dual-ballot
instructions or other highly individual forms of pre-electoral agreements are
rare. These figures may be read as a hint stronger commitment among political
parties who agree on nominations, present joint lists or present joint programs
because these cooperation forms entail additional transactions costs for them.

Does their investment yield any return on turnout? Figure
\ref{fig:boxplotTurnoutAndOurPecsInTillman} hints to first, purely exploratory
answers. Each boxplot in this graph visualizes the distribution shows the
residual within-variation in turnout after country- \textit{and} time-fixed
effects have been taken into account. This is variation left to explain for 
pre-electoral coalitions (see \cite{Mummolo.2018}).


hints to a first answers

```{r, boxplotTurnoutAndOurPecsInTillman, fig.cap="Mean-Deviated Turnout by PEC"}
rm(list = ls()[!(ls() %in% clean_workspace)])
# data objects
type_lookup <- c("No PEC" = 0, "Nomination agreement" = 1, "Joint list" = 2,
    "Dual-ballot Instructions" = 3, "Vote transfer Instructions" = 4,
    "Public commitment" = 5, "Other" = 6
)
pdta <- filter(country_panel, in_tillman == 1)
pdta[, "turnout_2way"] <- resid(
    lm(turnout ~ factor(iso3c) + factor(year2), data = pdta)
)
pdta <- pdta %>%
    mutate(any_type0 = pec_neu == 0) %>%
    select(c("election_id", "turnout_2way", paste0("any_type", 0:6), "any_prog", "any_incumbent")) %>%
    gather("key", "value", turnout_2way) %>%
    select(-key) %>%
    rename(turnout_2way = value) %>%
    gather("any_type", "value", any_type0:any_type6) %>%
    mutate(any_type_label = NA)
for (i in seq(type_lookup)) { # store full type label for each i in type_lookup
    mask <- which(
        pdta[["any_type"]] == paste0("any_type", type_lookup[i]) &
        pdta[["value"]] == 1
    )
    pdta[mask, "any_type_label"] <- names(type_lookup)[i]
}
pdta <- na.omit(pdta)  # Drop NA entries in any_type_label -> gather() artifacts
# Create data frame for each row in the final plot.
pdta_incumb <- rename(pdta, status = any_incumbent) %>%
    mutate(key_label = "Any incumbent?") %>%
    select(-any_prog)
pdta_prog <- rename(pdta, status = any_prog) %>%
    mutate(key_label = "Any joint program?") %>%
    select(-any_incumbent)
pdta <- mutate(pdta, status = 0, key_label = "Marginal") %>%
    select(-any_prog, -any_incumbent)
pdta <- rbind.data.frame(pdta, pdta_incumb, pdta_prog)  # Stack DFs for plot.
pdta <- mutate(pdta,  # Prettify plot labels
    any_type_label = factor(any_type_label, levels = c(
        "Public commitment", "Nomination agreement", "Joint list",
        "Dual-ballot Instructions", "Vote transfer Instructions", "Other",
        "No PEC"),
        labels = c("Public\nCommitment", "Nomination\nagreement", "Joint\nlist",
            "Dual-ballot\nInstructions", "Vote Transfer\nInstructions",
            "Other", "No PEC")
    ),
    key_label = factor(as.character(key_label),
        levels = c("Marginal", "Any incumbent?", "Any joint program?")),
    status_label = factor(status, levels = 0:1, labels = c("No", "Yes"))
)
ggplot(data = pdta, aes(x = any_type_label, y = turnout_2way,
    fill = status_label)
) +
    geom_boxplot(outlier.fill = "white", outlier.shape = 21) +
    scale_x_discrete(drop=FALSE) +
    scale_fill_brewer(type = "qual", palette = 4) +
    labs(x = "Type of PEC", y = "Within-Differences in Turnout\nTwo-way Fixed-Effects") +
    facet_grid(key_label ~ .) +
    ggthemes::theme_fivethirtyeight() +
    theme(
        axis.title = element_text(),
        legend.title = element_blank()
    )
# housekeeping
rm(list = ls()[!ls() %in% clean_workspace])
```


```{r, edaRelFreqVariantsPEC, fig.cap="Election Level Distribution of PEC Properties"}
rm(list = ls()[!(ls() %in% clean_workspace)])
pdta <- subset(country_panel,
    subset = (in_tillman == 1),
    select = c("iso3c", "election_id",
        names(country_panel)[grep("any_", names(country_panel))]
    )
)
pdta <- gather(pdta, "key", "value", grep("any_", names(pdta)))
pdta <- aggregate(value ~ key, data = pdta, FUN = mean, na.rm = TRUE)
pdta <- mutate(pdta,
    shape_label = factor(
        ifelse(as.character(key) %in% paste0("any_type", seq(6)),
            "PEC Type", "Other"
        ),
        levels = c("PEC Type", "Other")
    ),
    key_label = factor(key,
        levels = c(paste0("any_type", seq(6)), "any_prog", "any_incumbent"),
        labels = c("Nomination\nAgreement", "Joint list",
            "Dual-Ballot\nInstructions", "Vote Transfer\nInstructions",
            "Public commitment\nto govern together", "Other", "Joint Program",
            "Incumbent Coalition\nGovernment"
        )
    )
)
ggplot(data = pdta, aes(x = key_label, y = value, fill = shape_label)) +
    geom_bar(stat = "identity") +
    labs(y = "Relative Frequency", caption=paste("Reading instruction:",
            "In about 10 percent of all elections in the dataset at least one",
            "PEC formed a nomination agreement."
        )
    ) +
    scale_fill_brewer(type = "qual") +
    ggthemes::theme_fivethirtyeight() +
    theme(
        axis.title = element_text(), axis.title.x = element_blank(),
        legend.title = element_blank()
    )
rm(list = ls()[!(ls() %in% clean_workspace)])
```



```{r, tablePecsByType, eval = FALSE}
# This snippet returns the data required to reproduce table "XXX"
rm(list = ls()[!(ls() %in% clean_workspace)])

# functions
flag_nas <- function(r) all(is.na(r))

# data objects
pecs <- read_dta(
    file.path(path_project, "dta", "raw", "PEC_raw_stable15_2.dta"),
    encoding = "latin1"
)

# constants
pec_indicators <- grep("^[pec1-9]{4}$", names(pecs), perl = TRUE, value = TRUE)
suffixes <- c("type", "prog", "incumbent", "ea")
pec_indicators_2_aggregate <- character(0)
for (suffix in suffixes) {
    pec_indicators_2_aggregate <- append(
        pec_indicators_2_aggregate, paste(pec_indicators, suffix, sep = "_")
    )
}

# filter data
# (a) remove parties that did not enter PECs
pecs[, "no_pecs"] <- apply(pecs[, pec_indicators], 1, flag_nas)
pecs <- filter(pecs, !no_pecs)
# (b) aggregate to single pec entry per election
tmp <- subset(pecs, select = c("election_id", pec_indicators_2_aggregate))
tmp <- aggregate(
    x = as.data.frame(tmp[, -1]),
    by = list(election_id = tmp[["election_id"]]),
    FUN = function(x) ifelse(all(is.na(x)), NA, max(x, na.rm = TRUE))
)
# sanity check: n type codings <= nupec_neu
sum(!is.na(tmp[, paste(pec_indicators, "type", sep="_")])) <=
    sum(country_panel$nupec_neu, na.rm = TRUE)
# sanity check: n is.na(type) + n !(is.na(type)) == nupec_neu
sum(apply(tmp[, paste(pec_indicators, "type", sep="_")], 1, flag_nas)) +
    sum(!is.na(tmp[, paste(pec_indicators, "type", sep="_")])) ==
    sum(country_panel$nupec_neu, na.rm = TRUE)
# FAILS! 29 all NA entries, should be 12
tmp[apply(tmp[, paste(pec_indicators, "type", sep="_")], 1, flag_nas), 1]
# Elections in question: 20 166 171 205 227 284 291 306 357 386 403 407 453 479
# 496 508 596 606 623 643 653 664 709 710 711 712 713 715 782

# sanity checks: n codings <= sum of nupec_neu?
sum(!is.na(tmp[, paste(pec_indicators, "incumbent", sep="_")]))
sum(!is.na(tmp[, paste(pec_indicators, "prog", sep="_")]))
sum(!is.na(tmp[, paste(pec_indicators, "ea", sep="_")]))
# All pass

# sanity checks: n is.na(suffixes) + n !(is.na(suffixes)) == nupec_neu
sum(apply(tmp[, paste(pec_indicators, "incumbent", sep="_")], 1, flag_nas)) +
    sum(!is.na(tmp[, paste(pec_indicators, "incumbent", sep="_")])) # fails
sum(apply(tmp[, paste(pec_indicators, "prog", sep="_")], 1, flag_nas)) +
    sum(!is.na(tmp[, paste(pec_indicators, "prog", sep="_")])) # fails
sum(apply(tmp[, paste(pec_indicators, "ea", sep="_")], 1, flag_nas)) +
    sum(!is.na(tmp[, paste(pec_indicators, "ea", sep="_")])) # fails
# This test fails every time. Something about the test must be wrong

# create table
for (suffix in suffixes) {
    selection <- paste(pec_indicators, suffix, sep = "_")
    assign(suffix, gather(tmp[, selection], "pec", "value", selection)[[2]])
    # names(suffix, character.only = TRUE)
}
type <- factor(type, seq(6), c("Nomination agreement", "Joint list",
    "Dual-ballot instructions", "Vote transfer instructions", "Public commitment",
    "other"
    )
)
for (i in suffixes[-1]){
    assign(i, factor(get(i), 0:1, c("No", "Yes")))
}
lapply(mget(suffixes), function(x) cumsum(table(x)) / sum(table(x)))

# housekeeping
rm(list = ls()[!(ls() %in% clean_workspace)])
```

\begin{table}[htb]
    \centering
    \begin{tabular}{llccc}
    \toprule
    Attribute & Value & Freq. & Cum. Freq. & Cum. Rel. Freq.\\
    \midrule
    Type & Nomination agreement & 72 & 72 & 0.14\\
        ~ & Joint list & 204 & 276 & 0.63\\
        ~ & Dual-ballot instructions & 1 & 277 & 0.64\\
        ~ & Vote transfer instructions & 24 & 301 & 0.69\\
        ~ & Public commitment & 119 & 420 & 0.96\\
        ~ & Other & 16 & 436 & 1.00\\
    \midrule
    Joint program & No & 233 & 233 & 0.54\\
        ~ & Yes & 197 & 430 & 1.00\\
    \midrule
    Incumbent coalition & No & 337 & 337 & 0.77\\
        ~ & Yes & 99 & 436 & 1.00\\
    \bottomrule
    \end{tabular}
    \caption{Summary of pre-electoral coalition properties, 1945-2015}
    \label{tbl:SummaryPecProperties}
\end{table}


# Discussion

# Conclusion

# Appendix

## Country-level trends

```{r, edaCommonTrend, fig.cap="Country-Specific Trends in the Data"}
rm(list = ls()[!(ls() %in% clean_workspace)])
max_degree = 3
pdta <- select(tillman, country, year, turnout, pec1, vote_pec) %>%
    group_by(country) %>%
    mutate_at(vars(turnout, pec1, vote_pec), scale) %>%
    ungroup() %>%
    mutate(country = ifelse(country == "NewZealand", "New Zealand", country)) %>%
    gather("key", "value", turnout, pec1, vote_pec) %>%
    mutate(key = factor(as.character(key),
            levels = c("turnout", "pec1", "vote_pec"),
            labels = c("Turnout", "PEC contested", "PEC result")
        )
    )
ggplot(data = pdta, aes(x = year, y = value, linetype = key)) +
    geom_smooth(
        method = "lm", formula = y ~ poly(x, max_degree), se = FALSE, colour = "black", size = .7
    ) +
    geom_rug(sides = "b") +
    ggthemes::theme_fivethirtyeight(base_size = base_size * .75) +
    facet_wrap(~ country) +
    labs(
        y = "Scaled Within-Differences", linetype = "",
        caption = "Note: Figures scaled by within-country std. dev. Cubic polynomials shown. Rugs indicate elections."
    ) +
    theme(
        axis.title = element_text(),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "bottom"
    )
```

## Detailed Replications Results

\include{detrended_models}

## Coverage of our data on PECs

Following \cite{Golder.2006}, we regard as pre-electoral coalition any group of parties that publicly and mutually announces its intention to form or continue
a coalition government prior to election night. Obviously, such statements may
differ in their degree of commitment. Some coalitions do not even agree on
written terms whereas others present joint programs or field joint candidates.
We provide new and unprecedentedly granular data on
`r sprintf("%.0f", sum(country_panel$nupec_neu, na.rm=TRUE))` pre-electoral coalitions. Moreover, covering
`r sprintf("%.0f", length(unique(country_panel$election_id[!is.na(country_panel$in_lsvergl)])))` elections held between $1945$ and $2015$ in `r sprintf("%.0f", length(unique(country_panel$iso3c[!is.na(country_panel$in_lsvergl)])))` non-presidential democracies our data
provide the most complete survey of pre-electoral coalitions yet. Table \ref{tbl:summaryPecs} summarizes our data.

```{r, pecsSummaryTable, results = "asis"}
# EDA: Return basic data information table
rm(list = ls()[!(ls() %in% clean_workspace)])
packs <- c("lubridate", "stargazer")
missing <- which(!(packs %in% rownames(installed.packages())))
if (any(missing)) {
    cat("Installing missing packages: ", packs[missing], "\n")
    install.packages(packs[missing], dependencies = TRUE)
}
for (p in packs) library(p, character.only = TRUE, quietly = TRUE)

# Constants
col_names <- c("Country", "First election", "Last election",
    "# Elections", "# PECs", "Min", "Med", "Max"
)


# Data objects
tmp <- as_tibble(filter(country_panel, in_lsvergl == 1)) %>%
    mutate(edate = ymd(election_date)) %>%
    group_by(iso3c)
first_election <- summarize_at(tmp, vars(edate), min) %>%
    rename("first_election" = edate)
last_election <- summarize_at(tmp, vars(edate), max) %>%
    rename("last_election" = edate)
n_elections <- summarize_at(tmp, vars(edate), function(x) length(unique(x))) %>%
    rename("n_elections" = edate)
n_pecs <- summarize_at(tmp, vars(nupec_neu), sum) %>%
    rename("n_pecs" = nupec_neu)
range_n_pecs <- summarize_at(tmp, vars(nupec_neu),
    .funs = list(min = min, median = median, max = max)
)
tmp <- plyr::join_all(
    list(first_election, last_election, n_elections, n_pecs, range_n_pecs),
    by = "iso3c", type = "inner"
)
tmp <- tmp[with(tmp, order(n_pecs)), ]
names(tmp) <- col_names


# Return table
stargazer(
    tmp,
    title = "Summary of pre-electoral coalitions and elections by country",
    label = "tbl:summaryPecs",
    summary = FALSE, style = "apsr", header = FALSE, rownames = FALSE
)


# housekeeping
for (p in packs) detach(paste("package", p, sep = ":"), character.only = TRUE)
rm(list = ls()[!(ls() %in% clean_workspace)])
```

For each country in our data, Table \ref{tbl:summaryPecs} reports the total number
of elections and pre-electoral coalitions included as well as for each the
minimum,  median, and maximum count of contending PECs.


## Trends in PECs

```{r, edaPecOverTime, fig.cap="Distribution of PECs over Time"}
rm(list = ls()[!(ls() %in% clean_workspace)])


# Prep plot data
nupec_per_year <- aggregate( # How many PECs per year?
    nupec_neu ~ year,
    data = subset(country_panel, in_lsvergl == 1), FUN = sum
)
nelec_per_year <- aggregate(  # How many elections per year?
    election_id ~ year, data = subset(country_panel, in_lsvergl == 1),
    FUN = function(x) length(x)
)
perc_pec_per_year <- filter(country_panel, in_lsvergl == 1) %>%
    # relative freq of PECs in elections per year
    select(year, nupec_neu) %>%
    mutate(pec = ifelse(nupec_neu > 0, 1, 0)) %>%
    group_by(year) %>%
    summarise_at(vars(pec), mean)
pdta <- plyr::join_all(list(nupec_per_year, nelec_per_year, perc_pec_per_year),
    by = "year", type = "inner"
)
pdta <- gather(pdta, "key", "value", nupec_neu, election_id, pec)
pdta <- within(pdta,
    key_labels <- factor(as.character(key),
        levels = c("nupec_neu", "election_id", "pec"),
        labels = c("No. of PECs", "No. of elections",
            "Share elections with PECs"
        )
    )
)


# Return plot
ggplot(data = pdta, aes(x = year, y = value)) +
    geom_bar(stat = "identity", width = .9) +
    facet_grid(key_labels ~ ., scales = "free_y") +
    scale_x_continuous(breaks = seq(1945, 2015, 10)) +
    labs(y = "Frequency") +
    theme(axis.title.x = element_blank())

# housekeeping
rm(list = ls()[!(ls() %in% clean_workspace)])
```

