---
output:
    pdf_document:
        includes:
            in_header: header.tex
    documentclass: cup_PSRM
---

\markboth{GANGHOF, EPPNER, AND TANNEBERG}{Turnout and Pre-Electoral Coalitions}

```{r, optionsChunk, include = FALSE}
knitr::opts_chunk$set(
    cache = FALSE, eval = TRUE, echo = FALSE, fig.align = "center",
    fig.width = 12, fig.height = 8, warning = FALSE, message = FALSE
)
source("~/github/pecs/src/r/master.R")
```

<!-- \journalname{Draft Submission to Political Science Research and Methods}
\journalcopy{The European Political Science Association, 2019}
\fpage{X}
\lpage{XXX}
\journalvolume{6}
\journalissue{3}
\doinumber{doi:10.1017/psrm.2016.XX}
 -->

\author{Steffen Ganghof, Sebastian Eppner, Dag Tanneberg}
\title{Bring out the Rest\newline What pre-electoral coalitions increase turnout in national elections?}
\maketitle

# Introduction

What explains turnout in general elections with high numbers of competing parties?
After all, as the number of political parties increases coalition cabinets become
more likely whose format usually depends exclusively on the bargains struck in
parliament. ``The presence of many parties may mean that voters have little say in
the actual selection of the government'' \cite[118]{Blais.2006}, effectively
reducing the incentive to turn out and vote. Extant research suggests that
pre-electoral coalitions mitigate the adverse effect of multipartism as they
provide everyone with additional information on credible coalition options
\cite[197]{Smith.2018}. However, prior empirical tests of that argument
suffer from two non-negligble constraints. First, results are based on a small set
of countries and a limited number of elections. As a consequence, their
generalizability is questionable. Second, and more importantly, prior research
frequently aggregates data to the election level, disregarding information on
different types, electoral strength or prior incumbency of pre-electoral
coalitions. This paper presents new data on pre-electoral coalitions that have
higher temporal and spatial coverage and that allow closer scrutiny of each
covered pre-electoral coalition. The data highlight substantial differences
between pre-electoral coalitions and imply that only coalitions which formulate
a clear alternative to the incumbent government increase turnout in multiparty
elections.

# Why does multipartism decrease turnout?

# What pre-electoral coalitions increase government identifiability?

# Introducing the data

Following \cite{Golder.2006}, we regard as pre-electoral coalition any group of
parties that publicly and mutually announces its intention to form or continue
a coalition government prior to election night. Obviously, such coalitions
differ in their degree of formalization. Participating parties rarely agree on
written terms or field joint candidates. In fact, we content that such
differences between pre-electoral coaltions are more informative of
within-coalition dynamics than anything else. Notwithstanding, parties usually
honor pre-electoral coalitions no matter their level of formalization
\cite[728]{Tillman.2015}. We provide new data in unprecedented granularity on
`r sprintf("%.0f", sum(country_panel$nupec_neu, na.rm=TRUE))` pre-electoral coalitions that contested elections in `r sprintf("%.0f", length(unique(country_panel$iso3c[!is.na(country_panel$in_lsvergl)])))` non-presidential democracies between 1945 and
2015. Table \ref{tbl:summaryPecs} summarizes our coverage.

```{r, pecsSummaryTable, results = "asis"}
# Create a simple eda table
# Preamble
rm(list = ls()[!(ls() %in% clean_workspace)])
packs <- c("lubridate", "stargazer")
missing <- which(!(packs %in% rownames(installed.packages())))
if (any(missing)) {
    cat("Installing missing packages: ", packs[missing], "\n")
    install.packages(packs[missing], dependencies = TRUE)
}
for (p in packs) library(p, character.only = TRUE, quietly = TRUE)

# Constants
col_names <- c("Country", "First election", "Last election",
    "No. of elections", "No. of PECs"
)


# Data objects
tmp <- as_tibble(filter(country_panel, in_lsvergl == 1)) %>%
    mutate(edate = ymd(election_date)) %>%
    group_by(iso3c)
first_election <- summarize_at(tmp, vars(edate), min) %>%
    rename("first_election" = edate)
last_election <- summarize_at(tmp, vars(edate), max) %>%
    rename("last_election" = edate)
n_elections <- summarize_at(tmp, vars(edate), function(x) length(unique(x))) %>%
    rename("n_elections" = edate)
n_pecs <- summarize_at(tmp, vars(nupec_neu), sum) %>%
    rename("n_pecs" = nupec_neu)
tmp <- plyr::join_all(
    list(first_election, last_election, n_elections, n_pecs),
    by = "iso3c", type = "inner"
)
tmp <- tmp[with(tmp, order(n_pecs)), ]
names(tmp) <- col_names

# Return table
stargazer(
    tmp,
    title = "Summary of pre-electoral coalitions and elections by country",
    label = "tbl:summaryPecs",
    summary = FALSE, style = "apsr", header = FALSE, rownames = FALSE
)


# housekeeping
for (p in packs) detach(paste("package", p, sep = ":"), character.only = TRUE)
rm(list = ls()[!(ls() %in% clean_workspace)])
```



# Results

# Discussion

# Conclusion
